`phyr`: An R package for phylogenetic species-distribution modelling in ecological communities

### 2020-02-13 ms.Rmd
update ms [skip ci]

---
title: "`phyr`: An R package for phylogenetic species-distribution modelling in ecological communities"
author: "Daijiang Li, Russell Dinnage, Lucas Nell, Matthew Helmus, Anthony Ives"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::word_document2: 
    toc: no
  bookdown::html_document2:
    number_sections: no
    theme: sandstone
    toc: yes
    css: style.css
    keep_md: no
  bookdown::tufte_html2:
    number_sections: no
    toc: yes
  redoc::redoc:
fontsize: 12pt
link-citations: yes
always_allow_html: yes
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/methods-in-ecology-and-evolution.csl
bibliography: ref.bib
---

```{r knitr_options, include = FALSE}
if(!require("xfun")) install.packages("xfun")
if(!require("phyr")) devtools::install_github("daijiang/phyr")
xfun::pkg_attach2(c("phyr", "knitr", "kableExtra", "kableExtra", "tidyverse"))
# opts_knit$set(root.dir = normalizePath("../"))
opts_chunk$set(fig.align='center', cache = TRUE, warning = FALSE, 
               message = FALSE, echo = TRUE)
options(digits = 3, width = 88, knitr.kable.NA = '')
# traits, phy, dat
source(file = "simulation_pglmm.R", echo = FALSE, print.eval = FALSE)
```

@@

@@

1. Model-based approaches are increasingly popular in ecological studies. A good example of this trend is the use of joint species distribution models to ask questions about ecological communities. However, most current applications of model-based methods do not include phylogenies despite the well-known importance of phylogenetic relationships in shaping species distributions and community composition. In part, this is due to lack of accessible tools allowing ecologists to fit phylogenetic species distribution models easily.
2. To fill this gap, the R package `phyr` (pronounced fire) implements a suite of metrics, comparative methods and mixed models that use phylogenies to understand and predict community composition and other ecological and evolutionary phenomena. The `phyr` workhorse functions are implemented in C++ making all calculations and model estimations fast.
3. `phyr` can fit a variety of models such as phylogenetic joint-species distribution models, spatiotemporal-phylogenetic autocorrelation models, and phylogenetic trait‐based bipartite network models. `phyr` also estimates phylogenetically independent trait correlations with measurement error to test for adaptive syndromes and performs fast calculations of common alpha and beta phylogenetic diversity metrics. All `phyr` methods are united under Brownian motion or Ornstein-Uhlenbeck models of evolution and the inclusion of phylogenetic terms based on phylogenetic covariance matrices.
4. The functions and model formula syntax we propose in `phyr` serves as a simple and unified framework that ignites the use of phylogenies to address a variety of ecological questions.

@@

(ref:maunFuncCaption) List of main functions in `phyr` package.

```{r mainFunc, echo=FALSE}
read.csv("phyr_func.csv", check.names = FALSE) %>% 
  knitr::kable(caption = "(ref:maunFuncCaption)") #%>%
  # kableExtra::kable_styling(bootstrap_options = c("striped", "scale_down", "hover"),
  #               latex_options = c("striped")) %>% 
  # kableExtra::column_spec(2, width = "25%")
```

@@

`pglmm()` constructs and fits generalized linear mixed models that incorporate covariance matrices containing the phylogenetic relationships among species. The syntax for `pglmm()` resembles that used in the R package `lme4` [@bates2015lme4], and indeed `pglmm()` will fit most of the models that can be fit with `lmer()` and `glmer()`. `pglmm()` goes beyond `lmer()`and `glmer()` by allowing the specification of covariance matrices, which could be phylogenetic covariance matrices or any other covariance matrices that the user defines (e.g., spatial or temporal autocorrelation matrix). `pglmm()` can also fit models with “nested” covariance structures (e.g., a species phylogenetic covariance matrix nested within a site covariance matrix). `pglmm()` can operate in both frequentist mode, with the distribution of species among communities being Gaussian, binary, binomial or Poisson, and Bayesian mode with the addition of zero-inflated binomial and Poisson distributions. Finally, it is our hope that the formula syntax of `pglmm()` can be used to fit similar models with other programs such as Stan [e.g. via R package `brms` @brmsR2018].

@@

Here, Y is a binary (Bernoulli) dependent variable which takes values of either 0 or 1. The specification `family = 'binomial'` allows binary data and also binomial data for which Y is a matrix containing columns for successes and failures. The independent variables `trait` and `env` take on different values for each species and site, respectively. Sites and species (`sp`) are treated as random effects: `(1|site)` implies that a value from a Gaussian random variable is picked for each site, thereby representing unmeasured differences among sites. For the case of species, the double underscore in `(1|sp__)` implies that, in addition to a random effect for species, there is a second random effect which contains the phylogenetic relationships among species. The phylogenetic random effect assumes that values for each species are picked from a multivariate Gaussian distribution with phylogenetic covariance matrix 𝚺. A covariance matrix 𝚺 is specified by `cov_ranef = list(sp = phy.sp, site = V.space)`. The covariance matrix `phy.sp` associated with species can be a `phylo` object from the R package `ape` [@ape2018]. It is also possible to specify an explicit covariance matrix, such as `site = V.space`. To construct 𝚺 from a 'phylo' object, `pglmm()` assumes that the residual variation associated with species follows a Brownian motion model of evolution, so that the covariance between species is proportional to their shared evolutionary history (e.g., shared branch length on a phylogeny).

The syntax `(1|sp__)` or  `(1|site__)` generates two random effects, one without and one with phylogenetic or spatial covariances; in contrast, `(1|sp)` would generate only a single random effect that is independent among species. `pglmm()` forces in a term for `(1|sp)` whenever `(1|sp__)` is specified, because otherwise any difference among species would be captured by the diagonal elements in 𝚺 even in the absence of covariances among phylogenetically related species which are specified by the off-diagonal elements of 𝚺. Therefore, if `(1|sp)` were not included, this could lead to the identification of phylogenetic signal in the abundances of species even in its absence. To account for differences among sites in how they select for species with different traits, `(trait|site)` allows the slope of Y against `trait` to be a Gaussian random variable. Similarly, to account for the differences among species for how they respond to `env`, `(env|sp__)` allows the relationship of Y against `env` to be given by two slopes, the first slope that is picked from a Gaussian random variable in which species are independent and the second slope that is picked from a multivariate Gaussian with covariance matrix 𝚺. Finally, `(1|sp__@site)` generates a nested term: within a site, the residual variation in Y shows phylogenetic relatedness, with phylogenetically related species more likely to occur in the same site. Note that `(1|sp__)` differs from `(1|sp__@site)` because `(1|sp__)` generates differences in the mean value of Y for species across all sites, whereas `(1|sp__@site)` is local to sites, giving the covariances among species only within sites. This nested term can be used to test for community clustering or overdispersion [@webb2002phylogenies; @ives2011generalized]. Other forms of a nested term are available in `pglmm()`, which can be used to study more complicated questions such as bipartite networks.

With `bayes = FALSE`, `pglmm()` is fitted using a frequentist approach. ML or REML is used for fitting, with `REML = TRUE` as the default. For a non-Gaussian model (e.g., `family = 'binomial'`), an iterated quasi-likelihood method is used for model fitting which gives the approximate likelihood; p-values for the fixed effects are given by a Wald test and for the random effects by profile likelihood, although we recommend bootstrap-based tests when computationally feasible. Note that `REML = TRUE` is an option for non-Gaussian models (in contrast to `glmer`) due to the algorithm used. With `bayes = TRUE`, a Bayesian approach is implemented using INLA [@rue2009approximate], which gives parameter estimates and credible intervals. For large problems with the number of species-site combinations exceeding 2000, the Bayesian computations are considerably faster than the frequentist computations. Finally, a key to interpreting the results from a model is understanding the structure of the covariance matrices associated with the random effects. Therefore, `pglmm()` has associated plotting functions `pglmm_plot_ranef()` that present the design matrices for the random effects (Fig. \@ref(fig:designPlot)).

@@

In this example, the correlation between `trait1` and `trait2` is computed, and the column named `sp` in `data` identifies the species. The object `phy.sp` specifies the phylogenetic covariance matrix as a 'phylo' object from the `ape` package. `cor_phylo()` estimates the phylogenetic signal for each trait by assuming that trait evolution is given by a Ornstein-Uhlenbeck process. The term `covariates = list(trait1 ~ env)` includes the independent variable `env` for `trait1`, to remove possible confounding effects; only an intercept is estimated if no covariate is provided for a trait. Within-species variation (measurement error) is specified by `meas_errors = list(trait1 ~ me1, trait2 ~ me2)`, where `me1` and `me2` are the standard errors for `trait1` and  `trait2`, respectively, of values at the tips of the phylogenetic tree. If within-species standard errors are not provided for a given trait, the trait values are assumed to be known without error. Finally, `cor_phylo()` can perform parametric bootstrapping to give confidence intervals for all parameter estimates: correlations, phylogenetic signals, covariate coefficients, and coefficient covariances.

@@

`pglmm()` is syntactically modeled after `lmer()` and `glmer()` in `lme4` [@bates2015lme4], although it allows the specification of phylogenetic covariance matrices. `pglmm()` also allows "nested" models (with block-diagonal covariance matrices) which arise when phylogenetic covariances only act within single communities, rather than among communities. Such nested models make it possible to assess whether phylogenetic relatedness affects the abundance of species within the same communities, such as whether competition between closely related species excludes one of the competitors from communities where the other is present. Nested models are structurally incompatible with the architecture of `lme4`.

@@

Of course, the goals of solid inference and prediction are not mutually exclusive. Good prediction requires avoiding overfitting, which can be facilitated by reducing the number of parameters in the model. In some systems, it may be possible to make better predictions using a simple phylogenetic model if phylogeny is a strong predictor of species covariance, or if many species are poorly sampled and thus estimating covariances between them individually results in higher prediction variance. Using phylogeny can help closely related species share statistical strength through phylogenetic partial pooling. Ultimately, it can be powerful to use the same kind of statistical framework to do both predictive and inferential work in ecology, because it allows ecologists to smoothly move between these two goals, both of which are important in modern ecology, and more easily and quickly draw mutual insights between them.

@@

We fitted a PGLMM that examined how a hypothetical functional trait, environmental gradient, and their interaction affect distributions of `r nspp` species across `r nsite` sites. We focused on abundance and used the default family of data distribution (Gaussian), but other distributions can also be specified by resetting the `family` argument. Phylogenetic relationships among species and site spatial autocorrelations are specified by `cov_ranef = list(species = phy, site = Vspace)` where `species` and `site` are group variables of random terms, `phy` can be a phylogeny with class `phylo` or a phylogenetic covariance matrix, `Vspace` is a variance-covriance matrix among sites. This model can also be fitted with Bayesian framework by setting `bayes = TRUE`, which is recommended when data set is large.

```{r, eval=FALSE}
z <- pglmm(
  abund ~ 1 + envi + trait + envi:trait +
    (1 | species__)  + (1 | site__) + 
    (envi | species__) + (trait | site) +
    (1 | species__@site),
  data = dat, 
  cov_ranef = list(species = phy, site = Vspace)
)
```

@@

The model formula of `pglmm()` is general and can be applied using other tools to fit phylogenetic species-distribution models. Thus, `pglmm()` can serve the developer community as a shell for new methods that fit GLMMs, with `phyr` providing an easy user interface. Using INLA as a backend to fit a Bayesian version of the pglmm model is an example of this approach. To facilitate this end, we are developing phyr openly on github and actively encourage community contribution.We hope that the `phyr` package will help current and future researchers formulate and analyze phylogenetic species-distribution models.

@@

Funding for this work was provided by the National Science Foundation (NSF ...).

@@

No data were used in this study. `phyr` is available at Github (https://github.com/daijiang/phyr) and CRAN (https://cran.r-project.org/package=phyr).

@@

To demonstrate the usage of main functions in phyr, we simulated a dataset with 30 species and 30 communities. We first simulated a coalescent phylogeny of 30 species with function `ape::rcoal()`. For each species, we then simulated one continuous functional trait along the phylogeny. We also simulated one environmental variable with all 30 communities located evenly along the gradient. Environmental variable, functional trait, and their interaction all determine the abundance of species over sites following the model below.

$$y=(\beta_{0}+b_{0\_sp}+b_{0\_site})+(\beta_{1}+b_{1})envi+(\beta_{2}+b_{2})trait+\beta_{3}envi*trait+b_{3}+e$$
$$b_{0\_{sp}}\sim Gaussian(0,\sigma_{0\_{sp}}^{2}\Sigma_{spp})$$
$$b_{0\_{site}}\sim Gaussian(0,\sigma_{0\_{site}}^{2}\Sigma_{site})$$

$$b_{1}\sim Gaussian(0,\sigma_{1}^{2}\Sigma_{spp})$$

$$b_{2}\sim Gaussian(0,\sigma_{2}^{2}I_{site})$$
$$b_{3}\sim Gaussian(0,\sigma_{3}^{2}\Sigma_{site}\otimes\Sigma_{spp})$$

$$e\sim Gaussian(0,\sigma_{e}^2)$$

Here, we set all coefficients ($\beta_0$ to $\beta_3$) to be 1; we also set variances of all random terms to be 1. $\Sigma_{spp}$ is a variance-covariance matrix converted from the phylogeny. $\Sigma_{site}$ is a vairance-covariance matrix among sites, which was converted from the site spatial distance matrix. $I_{site}$ is an identity matrix so that we treat sites as independent replications. $\Sigma_{site}\otimes\Sigma_{spp}$ a kronecker product and forces closely related species will be more likely to be observed in the same site (i.e. underdispersion). In the above model, different species have different overall abundance (intercept) but closely related species ($b_{0\_{sp}}$) and similar communities ($b_{0\_{site}}$) have similar overall abundance. Abundance of different species also change differently along the environmental gradient; however, closely related species again change similarly in response to environmental change. We included a random term for functional trait; therefore the relationship between functional traits and species abundance varied across sites independently to account for unmeasured environmental variables. For demonstration purposes, we only simulated one functional trait and one environmental variable; in real datasets, multiple environmental variables and multiple functional traits can be included in the model. However, more random terms will need more computational time to finish the model fitting.

>>>>

---
title: "`phyr`: An R package for phylogenetic species-distribution modelling in ecological communities"
author: "Daijiang Li^1^, Russell Dinnage^2^, Lucas Nell^3^, Matthew Helmus^4^, Anthony Ives^3^"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    number_sections: no
    theme: sandstone
    toc: yes
    css: style.css
    keep_md: no
  bookdown::word_document2: 
    toc: no
  bookdown::pdf_document2:
    includes:
      before_body: doc_prefix.tex
      in_header: preamble.tex
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
    toc: no
  bookdown::tufte_html2:
    number_sections: no
    toc: yes
  redoc::redoc:
fontsize: 12pt
link-citations: yes
always_allow_html: yes
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/methods-in-ecology-and-evolution.csl
bibliography: ref.bib
---

```{r knitr_options, include = FALSE}
if(!require("xfun")) install.packages("xfun")
if(!require("phyr")) devtools::install_github("daijiang/phyr")
xfun::pkg_attach2(c("phyr", "knitr", "kableExtra", "kableExtra", "tidyverse"))
# opts_knit$set(root.dir = normalizePath("../"))
opts_chunk$set(fig.align='center', cache = TRUE, warning = FALSE, 
               message = FALSE, echo = TRUE)
options(digits = 3, width = 88, knitr.kable.NA = '')
tpl_1 = 'https://raw.githubusercontent.com/daijiang/workflow_demo/master/template/preamble.tex'
tpl_2 = 'https://raw.githubusercontent.com/daijiang/workflow_demo/master/template/doc_prefix.tex'
if(!file.exists(tpl_1f <- 'preamble.tex')) download.file(tpl_1, tpl_1f)
if(!file.exists(tpl_2f <- 'doc_prefix.tex')) download.file(tpl_2, tpl_2f)
# traits, phy, dat
source(file = "simulation_pglmm.R", echo = FALSE, print.eval = FALSE)
```

@@

^1^ Department of Wildlife Ecology & Conservation, University of Florida, Gainesville, FL 32611  
^2^ Macroevolution & Macroecology Group, Research School of Biology, Australian National University, Canberra, ACT, Australia    
^3^ Department of Integrative Biology, University of Wisconsin-Madison, Madison, WI 53706  
^4^ Integrative Ecology Lab, Center for Biodiversity, Department of Biology, Temple University, Philadelphia, PA 19122

Emails: daijianglee@gmail.com; r.dinnage@gmail.com; lucas.nell@wisc.edu; mrhelmus@temple.edu; arives@wisc.edu

@@

1. Model-based approaches are increasingly popular in ecological studies. A good example of this trend is the use of joint species distribution models to ask questions about ecological communities. However, most current applications of model-based methods do not include phylogenies despite the well-known importance of phylogenetic relationships in shaping species distributions and community composition. In part, this is due to lack of accessible tools allowing ecologists to fit phylogenetic species distribution models easily.
2. To fill this gap, the R package `phyr` (pronounced fire) implements a suite of metrics, comparative methods and mixed models that use phylogenies to understand and predict community composition and other ecological and evolutionary phenomena. The `phyr` workhorse functions are implemented in C++ making all calculations and model estimations fast.
3. `phyr` can fit a variety of models such as phylogenetic joint-species distribution models, spatiotemporal-phylogenetic autocorrelation models, and phylogenetic trait‐based bipartite network models. `phyr` also estimates phylogenetically independent trait correlations with measurement error to test for adaptive syndromes and performs fast calculations of common alpha and beta phylogenetic diversity metrics. All `phyr` methods are united under Brownian motion or Ornstein-Uhlenbeck models of evolution and phylogenetic terms are modelled as phylogenetic covariance matrices.
4. The functions and model formula syntax we propose in `phyr` serves as a simple and unified framework that ignites the use of phylogenies to address a variety of ecological questions.

@@

(ref:mainFuncCap) List of main functions in the `phyr` package.

```{r mainFunc, echo=FALSE}
read.csv("phyr_func.csv", check.names = FALSE) %>% 
  knitr::kable(caption = "(ref:mainFuncCap)", booktabs = T) %>%
  kableExtra::kable_styling(latex_options = "striped", full_width = T) %>%
  kableExtra::column_spec(3, width = "22em") 
```

@@

Function `pglmm()` constructs and fits generalized linear mixed models that incorporate covariance matrices containing the phylogenetic relationships among species. The syntax for `pglmm()` resembles that used in the R package `lme4` [@bates2015lme4], and indeed `pglmm()` will fit most of the models that can be fit with `lmer()` and `glmer()`. `pglmm()` goes beyond `lmer()`and `glmer()` by allowing the specification of covariance matrices, which could be phylogenetic covariance matrices or any other covariance matrices that the user defines (e.g., spatial or temporal autocorrelation matrix). `pglmm()` can also fit models with “nested” covariance structures (e.g., a species phylogenetic covariance matrix nested within a site covariance matrix). `pglmm()` can operate in both frequentist mode, with the distribution of species among communities being Gaussian, binary, binomial or Poisson, and Bayesian mode with the addition of zero-inflated binomial and Poisson distributions. Finally, it is our hope that the formula syntax of `pglmm()` can be used to fit similar models with other programs such as Stan [e.g. via R package `brms` @brmsR2018].

@@

Here, Y is a binary (Bernoulli) dependent variable which takes values of either 0 or 1. The specification `family = 'binomial'` allows binary data and also binomial data for which Y is a matrix containing columns for successes and failures. The independent variables `trait` and `env` take on different values for each species and site, respectively. Sites and species (`sp`) are treated as random effects: `(1|site)` implies that a value from a Gaussian random variable is picked for each site, thereby representing unmeasured differences among sites. For the case of species, the double underscore in `(1|sp__)` implies that, in addition to a random effect for species, there is a second random effect which contains the phylogenetic relationships among species (or some other correlation structure specified by the user). The phylogenetic random effect assumes that values for each species are picked from a multivariate Gaussian distribution with phylogenetic covariance matrix $\Sigma$. A covariance matrix $\Sigma$ is specified by `cov_ranef = list(sp = phy.sp, site = V.space)`. The covariance matrix `phy.sp` associated with species can be a `phylo` object from the R package `ape` [@ape2018]. To construct $\Sigma$ from a 'phylo' object, `pglmm()` assumes that the residual variation associated with species follows a Brownian motion model of evolution, so that the covariance between species is proportional to their shared evolutionary history (e.g., shared branch length on a phylogeny). It is also possible to specify an explicit covariance matrix, such as `site = V.space`, where `V.space` is a covariance matrix created from the distance between sites.

The syntax `(1|sp__)` or  `(1|site__)` generates two random effects, one without and one with phylogenetic or spatial covariances; in contrast, `(1|sp)` would generate only a single random effect that is independent among species. `pglmm()` forces in a term for `(1|sp)` whenever `(1|sp__)` is specified, because otherwise any difference among species would be captured by the diagonal elements in $\Sigma$ even in the absence of covariances among phylogenetically related species which are specified by the off-diagonal elements of $\Sigma$. Therefore, if `(1|sp)` were not included, this could lead to the identification of phylogenetic signal in the abundances of species even in its absence. To account for differences among sites in how they select for species with different traits, `(trait|site)` allows the slope of Y against `trait` to be a Gaussian random variable. Similarly, to account for the differences among species for how they respond to `env`, `(env|sp__)` allows the relationship of Y against `env` to be given by two slopes, the first slope that is picked from a Gaussian random variable in which species are independent and the second slope that is picked from a multivariate Gaussian with covariance matrix $\Sigma$. Finally, `(1|sp__@site)` generates a nested term: within a site, the residual variation in Y shows phylogenetic relatedness, with phylogenetically related species more likely to occur in the same site. Note that `(1|sp__)` differs from `(1|sp__@site)` because `(1|sp__)` generates differences in the mean value of Y for species across all sites, whereas `(1|sp__@site)` is local to sites, giving the covariances among species only within sites. This nested term can be used to test for community clustering or overdispersion [@webb2002phylogenies; @ives2011generalized]. Other forms of a nested term are available in `pglmm()`, which can be used to study more complicated questions such as bipartite networks.

With `bayes = FALSE`, `pglmm()` is fitted using a frequentist approach. ML or REML is used for fitting, with `REML = TRUE` as the default. For a non-Gaussian model (e.g., `family = 'binomial'`), an iterated quasi-likelihood method is used for model fitting which gives the approximate likelihood; p-values for the fixed effects are given by a Wald test and for the random effects by profile likelihood, although we recommend bootstrap-based tests when computationally feasible. Note that `REML = TRUE` is an option for non-Gaussian models (in contrast to `glmer()`) due to the algorithm used. With `bayes = TRUE`, a Bayesian approach is implemented using INLA [@rue2009approximate], which gives parameter estimates and credible intervals. For large problems with the number of species-site combinations exceeding 2000, the Bayesian computations are considerably faster than the frequentist computations. Finally, a key to interpreting the results from a model is understanding the structure of the covariance matrices associated with the random effects. Therefore, `pglmm()` has associated plotting functions `pglmm_plot_ranef()` that present the design matrices for the random effects (Fig. \@ref(fig:designPlot)).

@@

In this example, the correlation between `trait1` and `trait2` is computed, and the column named `sp` in `data` identifies the species. The object `phy.sp` specifies the phylogenetic covariance matrix as a 'phylo' object from the `ape` package. `cor_phylo()` estimates the phylogenetic signal for each trait by assuming that trait evolution is given by a Ornstein-Uhlenbeck process. The term `covariates = list(trait1 ~ env)` includes the independent variable `env` for `trait1`, to remove possible confounding effects; only an intercept is estimated if no covariate is provided for a trait. Within-species variation is specified by `meas_errors = list(trait1 ~ me1, trait2 ~ me2)`, where `me1` and `me2` are the standard errors for `trait1` and  `trait2`, respectively, of values at the tips of the phylogenetic tree. If within-species standard errors are not provided for a given trait, the trait values are assumed to be known without error. Finally, `cor_phylo()` can perform parametric bootstrapping to give confidence intervals for all parameter estimates: correlations, phylogenetic signals, covariate coefficients, and coefficient covariances.

@@

`pglmm()` is syntactically modeled after `lmer()` and `glmer()` in `lme4` [@bates2015lme4], although it allows the specification of phylogenetic covariance matrices. `pglmm()` also allows "nested" models (with block-diagonal covariance matrices) which arise when phylogenetic covariances only act within single communities, rather than among communities; an example is illustrated by the `(1|sp__@site)` term in Fig. \@ref(fig:designPlot). Such nested models make it possible to assess whether phylogenetic relatedness affects the abundance of species within the same communities, such as whether competition between closely related species excludes one of the competitors from communities where the other is present. Nested models are structurally incompatible with the architecture of `lme4`.

@@

Of course, the goals of solid inference and prediction are not mutually exclusive. Good prediction requires avoiding overfitting, which can be facilitated by reducing the number of parameters in the model. In some systems, it may be possible to make better predictions using a simple phylogenetic model if phylogeny is a strong predictor of species covariance, or if many species are poorly sampled and thus estimating covariances between them individually results in higher prediction variance. Using phylogeny can help closely related species share statistical strength through phylogenetic partial pooling. Ultimately, it can be powerful to use the same kind of statistical framework to do both predictive and inferential work in ecology, because it allows ecologists to smoothly move between these two goals and more easily and quickly draw mutual insights between them.

@@

We fitted a PGLMM that examined how a hypothetical functional trait, environmental gradient, and their interaction affect distributions of `r nspp` species across `r nsite` sites. We focused on abundance and used the default family of data distribution (Gaussian), but other distributions can also be specified by resetting the `family` argument. Phylogenetic relationships among species and site spatial autocorrelations are specified by `cov_ranef = list(species = phy, site = V.space)` where `species` and `site` are group variables of random terms, `phy` can be a phylogeny with class `phylo` or a phylogenetic covariance matrix, `Vspace` is a variance-covriance matrix among sites. This model can also be fitted with Bayesian framework by setting `bayes = TRUE`, which is recommended when data set is large.

```{r, eval=FALSE}
z <- pglmm(
  abund ~ 1 + env + trait + env:trait +
    (1 | species__)  + (1 | site__) + 
    (env | species__) + (trait | site) +
    (1 | species__@site),
  data = dat, 
  cov_ranef = list(species = phy, site = V.space)
)
```

@@

The model formula of `pglmm()` is general and can be applied using other tools to fit phylogenetic species-distribution models. Thus, `pglmm()` can serve the developer community as a shell for new methods that fit GLMMs, with `phyr` providing an easy user interface. Using INLA as a backend to fit a Bayesian version of the PGLMM model is an example of this approach. To facilitate this end, we are developing phyr openly on github and actively encourage community contribution. We hope that the `phyr` package will help current and future researchers formulate and analyze phylogenetic species-distribution models.

@@

Funding for this work was provided by the National Science Foundation (US-NSF-DEB Dimensions of Biodiversity, 1240804).

@@

No data were used in this study. R code used to simulate data for PGLMMs as described in the Appendix is available at <https://github.com/daijiang/phyr_ms/blob/master/simulation_pglmm.R>. R code used to simulate data for comparative methods is available at <https://github.com/daijiang/phyr_ms/blob/master/simulation_cor_phylo.R>. `phyr` is available at Github (https://github.com/daijiang/phyr) and CRAN (https://cran.r-project.org/package=phyr).

\clearpage

@@

To demonstrate the usage of main functions in `phyr`, we simulated a dataset with 30 species and 30 communities. We first simulated a coalescent phylogeny of 30 species with function `ape::rcoal()`. For each species, we then simulated one continuous functional trait along the phylogeny. We also simulated one environmental variable with all 30 communities located evenly along the gradient. The environmental variable, functional trait, and their interaction all determine the abundance of species among sites according to the model below:

$$y=(\beta_{0}+b_{0\_sp}+b_{0\_site})+(\beta_{1}+b_{1})envi+(\beta_{2}+b_{2})trait+\beta_{3}envi*trait+b_{3}+e$$
$$b_{0\_{sp}}\sim Gaussian(0,\sigma_{0\_{sp}}^{2}\Sigma_{spp})$$
$$b_{0\_{site}}\sim Gaussian(0,\sigma_{0\_{site}}^{2}\Sigma_{site})$$
$$b_{1}\sim Gaussian(0,\sigma_{1}^{2}\Sigma_{spp})$$
$$b_{2}\sim Gaussian(0,\sigma_{2}^{2}I_{site})$$
$$b_{3}\sim Gaussian(0,\sigma_{3}^{2}\Sigma_{site}\otimes\Sigma_{spp})$$
$$e\sim Gaussian(0,\sigma_{e}^2)$$

We set all coefficients ($\beta_0$ to $\beta_3$) to 1; we also set variances of all random terms to 1. $\Sigma_{spp}$ is a covariance matrix converted from the phylogeny. $\Sigma_{site}$ is a covariance matrix among sites, which was converted from the site spatial distance matrix. $I_{site}$ is the identity matrix so that we treat sites as independent replicates. $\Sigma_{site}\otimes\Sigma_{spp}$ is a matrix generated as a kronecker product, and it makes closely related species more likely to be observed in the same site (i.e. underdispersion). Different species have different overall abundance (intercept) but closely related species ($b_{0\_{sp}}$) and similar communities ($b_{0\_{site}}$) have similar overall abundance. Abundances of different species also change differently along the environmental gradient; however, closely related species change similarly in response to the environmental variable. We included a random term for functional trait; therefore, the relationship between functional trait and species abundance varied across sites independently to account for unmeasured environmental variables. For demonstration purposes, we only simulated one functional trait and one environmental variable; in real datasets, multiple environmental variables and multiple functional traits can be included in the model.

\clearpage
